Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-11-28 contents/draft.tex
Update minor
* Introduction:   - previously -> existing   - Further research is required in the future와 다음 문장을 합침 * NaSH2015: CRF++, Lattice-based HMM * Overview of Recent Korean Morphological Analysis Methods   - Add NaSH2015 * Deep Learning and Tokenization:   - Split 2 subsections: Recent Deep Learning Approaches, Integrating Dictionary-based and Deep Learning Approaches * Recent Deep Learning Approaches   - 기존 딥러닝 연구와의 차별점 기술 * Integrating Dictionary-based and Deep Learning Approaches   - 서론에서의 전개와 일맥상통하도록 보강

    Against this background, this study considers how the dictionary-based morphological analysis method used by MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis in tokenizers, which is an essential preprocessing tool for deep learning, can be effectively improved, and a method is proposed.
    The dictionary-based morphological analysis method~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018} trained by the CRF (Conditional Random Fields) method~\cite{Lafferty2001} lists the candidate morphemes in the dictionary from a given sentence to form a lattice structure connected by a directed graph and determines the optimal morphological analysis path within it.
    The process of determining the optimal path in the lattice uses the Viterbi algorithm~\cite{Viterbi1967}, which determines the path that minimizes the cost of each morpheme node and the sum of the neighborhood costs of two consecutive morphemes.
    The main types of errors in these dictionary-based morphological analysis methods occur when new words that are not in the dictionary are used in a sentence, or when the optimal path calculation selects the incorrect result owing to bias.
    For example, it may be cost-effective to select one long morpheme than several short morphemes, but this may often lead to an incorrect analysis.
    The main motivation for this study was that the path that minimizes the costs for the nodes and links may not be the optimal path.

@@

    To identify cases in which a suboptimal solution is actually the best solution according to the best path calculation, we modified the best path calculation method to generate suboptimal analysis results and verified the extent to which they are correct.
    Although there are numerous different approaches to select the next-best path, we used the method of replacing a morpheme node on the optimal path with a lower-ranked node.
    As shown in Table~\ref{tab:maximum-performance}, we confirm the extent to which the analysis performance can be improved by replacing the optimal path with a lower-ranked node.
    We can consider the problem of finding the correct answer among the generated sub-optimals, similar to the problem of re-ranking search results in information retrieval~\cite{BaeYJ2021}.
    In ~\cite{ChoiYS2018}, the N-best analysis results generated by the seq2seq model were re-ranked based on a convolutional neural network to improve the performance.
    In this study, re-ranking was performed using two BERT models of different types and forms, as proposed in ~\cite{Nogueira2019}.
    Experimental results show that first-stage re-ranking improves the performance by over 20\% over previously written and spoken models, and second-stage re-ranking with a different type of input and a different type of pre-trained model further improves the performance by more than 30\% over previously written and spoken models.

    With this method, the performance of the dictionary-based morphological analysis method could be further improved; however, the overall analysis time increased when the morphological analysis system was configured, including the re-ranking model itself.
    However, it is feasible to use the results of multiple reranked morpheme analyses to update the connection costs between morphemes in a dictionary, similar to the backpropagation process in a typical neural network.
    It is also expected that the morphological analysis system with improved connection costs will be able to generate better re-ranking candidates, which will further improve performance by doing so iteratively.
    Further research is required in the future.
    In this study, only performance improvement using the second-stage re-ranking was covered as the scope of the study.
    The main contributions of this study are as follows.
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We explore the possibility of performance improvement by introducing a method to replace the optimal path with a suboptimal node and propose a method to effectively improve the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To improve the performance of dictionary-based analysis by re-ranking the morphological analysis results, we propose extending the performance improvement using different BERT models to perform two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We propose a new method for updating dictionary connection costs based on re-ranked morphological analysis results. We also outline directions for future research, suggesting potential improvements.
    \end{enumerate}
    These contributions provide important insights into the performance improvement of Korean morphological analysis and the direction of future research, and will serve as a useful reference for future researchers.

@@

    \begin{table*}[]
        \caption{Comparison of performance differences with previous studies}\label{tab:differences-with-previous-studies}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}ccc|cc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{Authors}              & \multirow{2}{*}{Model}                     & \multirow{2}{*}{Data (train, test)}         & \multicolumn{2}{c}{Performance}                           \\
            \cmidrule{4-5}
            ~                                     & ~                                          & ~                                           & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            Na, 2015~\cite{NaSH2015}              & CRF++, HMM                                 & Sejong 200k, 50k sentences                  & 95.22                      & 97.21                        \\
            Lee et al., 2016~\cite{LeeCH2016}     & Structural SVM                             & Sejong 666k, 74k eojeols                    & 96.41                      & -                            \\
            Li et al., 2017~\cite{Li2017}         & Seq2seq (GRU-based)                        & Sejong 90k, 10k sentences                   & 95.33                      & 97.15                        \\
            Na and Kim, 2018~\cite{NaSH2018}      & Lattice + HMM                              & Sejong 200k, 50k sentences                  & 96.35                      & 97.74                        \\
            Min et al., 2019~\cite{MinJW2019}     & Seq2seq (Transition-based)                 & Sejong 200k, 50k sentences                  & 96.34                      & 97.68                        \\
            Song and Park, 2019~\cite{SongHJ2019} & Seq2seq (BiLSTM-based)                     & Sejong 200k, 50k sentences                  & 95.68                      & 97.43                        \\
            Youn et al, 2021~\cite{YounJY2021}    & Seq2seq (BERT-based)                       & Sejong 675k, 75k sentences                  & 95.99                      & 97.94                        \\
            Shin et al, 2023~\cite{ShinHJ2023}    & Transformer(En) + BiLSTM                   & Sejong 769k, 87k sentences                  & 96.12                      & 97.74                        \\
            \midrule
            Proposed (without rerank)             & \multirow{3}{*}{Lattice + Transformer(En)} & \multirow{3}{*}{Sejong 194k, 10k sentences} & 95.23                      & 97.08                        \\
            Proposed (1-stage rerank)             & ~                                          & ~                                           & 96.63                      & 97.84                        \\
            Proposed (2-stage rerank)             & ~                                          & ~                                           & \textbf{96.87}             & \textbf{98.01}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

    Korean morphological analyses have seen an influx of various methodologies in recent years~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    The nature of the Korean language, being agglutinative, introduces challenges that have propelled researchers to develop inventive solutions, many of which have laid the groundwork for future research.

@@

    \subsection{Syllable-unit Morphological Analysis}\label{subsec:syllable-unit-morphological-analysis}
    To overcome the limitations of dictionary dependence, syllable-by-syllable morphological analysis has emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    This method either tags each syllable and then applies a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}.
    One notable drawback is the challenge of accurately pinpointing morpheme boundaries.
    Furthermore, as the sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Deep Learning and Tokenization}\label{subsec:deep-learning-and-tokenization}
    The advent of deep learning has reshaped the Korean morphological analysis landscape.
    High-performance deep learning models in NLP employ tokenization, dividing the text into smaller units and then converting these tokens into vectors as an input to the computational model~\cite{Mikolov2013}.
    The attempt to conduct subword tokenization, incorporating separate morphemes to cater to Korean linguistic characteristics is particularly significant.
    Incorporating accurate morpheme analysis results into tokenization can improve overall analysis performance.

    With advancements in technology and access to large datasets, the potential to integrate dictionary-based methods with deep learning techniques has emerged.
    As motivated in our study, MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis, utilized dictionary-based morphological analysis trained by the CRF method~\cite{Lafferty2001}.
    This method attempts to find the optimal morphological analysis path by forming a lattice structure as in ~\cite{Kudo2004, NaSH2014, NaSH2015, NaSH2018}.
    The amalgamation of dictionary-based methods with deep learning, as explored in our research, signifies the latest strides in this journey aimed at harnessing the strengths of both approaches.
    Future research avenues include refining these hybrid methods and exploring the capabilities of end-to-end models.

>>>>

    Against this background, this study considers how the dictionary-based morphological analysis method used by MeCab~\cite{MeCab}, an open software for Korean and Japanese morphological analysis in tokenizers, which is an essential preprocessing tool for deep learning, can be effectively improved, and a method is proposed.
    The dictionary-based morphological analysis method~\cite{Kudo2004, NaSH2014, NaSH2018} trained by the CRF (Conditional Random Fields) method~\cite{Lafferty2001} lists the candidate morphemes in the dictionary from a given sentence to form a lattice structure connected by a directed graph and determines the optimal morphological analysis path within it.
    The process of determining the optimal path in the lattice uses the Viterbi algorithm~\cite{Viterbi1967}, which determines the path that minimizes the cost of each morpheme node and the sum of the neighborhood costs of two consecutive morphemes.
    The main types of errors in these dictionary-based morphological analysis methods occur when new words that are not in the dictionary are used in a sentence, or when the optimal path calculation selects the incorrect result owing to bias.
    For example, it may be cost-effective to select one long morpheme than several short morphemes, but this may often lead to an incorrect analysis.
    The main motivation for this study was that the path that minimizes the costs for the nodes and links may not be the optimal path.

@@

    To identify cases in which a suboptimal solution is actually the best solution according to the best path calculation, we modified the best path calculation method to generate suboptimal analysis results and verified the extent to which they are correct.
    Although there are numerous different approaches to select the next-best path, we used the method of replacing a morpheme node on the optimal path with a lower-ranked node.
    As shown in Table~\ref{tab:maximum-performance}, we confirm the extent to which the analysis performance can be improved by replacing the optimal path with a lower-ranked node.
    We can consider the problem of finding the correct answer among the generated sub-optimals, similar to the problem of re-ranking search results in information retrieval~\cite{BaeYJ2021}.
    In ~\cite{ChoiYS2018}, the N-best analysis results generated by the seq2seq model were re-ranked based on a convolutional neural network to improve the performance.
    In this study, re-ranking was performed using two BERT models of different types and forms, as proposed in ~\cite{Nogueira2019}.
    Experimental results show that first-stage re-ranking improves the performance by over 20\% over existing written and spoken models, and second-stage re-ranking with a different type of input and a different type of pre-trained model further improves the performance by more than 30\% over existing written and spoken models.

    With this method, the performance of the dictionary-based morphological analysis method could be further improved; however, the overall analysis time increased when the morphological analysis system was configured, including the re-ranking model itself.
    However, it is feasible to use the results of multiple reranked morpheme analyses to update the connection costs between morphemes in a dictionary, similar to the backpropagation process in a typical neural network.
    It is also expected that the morphological analysis system with improved connection costs will be able to generate better re-ranking candidates, which will further improve performance by doing so iteratively.
    Further research is needed on this, and this study was limited to performance improvements with two-stage reranking.

    The main contributions of this study are as follows.
    \begin{enumerate}
        \item \textbf{Further improvement of dictionary-based morphological analysis method using suboptimal analysis results}: We explore the possibility of performance improvement by introducing a method to replace the optimal path with a suboptimal node and propose a method to effectively improve the dictionary-based morphological analysis method through deep learning.
        \item \textbf{Extending the performance improvement by introducing a two-stage re-ranking model}: To improve the performance of dictionary-based analysis by re-ranking the morphological analysis results, we propose extending the performance improvement using different BERT models to perform two rounds of re-ranking.
        \item \textbf{A method for updating connection costs in the dictionary and suggestions for future research}: We propose a new method for updating dictionary connection costs based on re-ranked morphological analysis results. We also outline directions for future research, suggesting potential improvements.
    \end{enumerate}
    These contributions provide important insights into the performance improvement of Korean morphological analysis and the direction of future research, and will serve as a useful reference for future researchers.

@@

    \begin{table*}[]
        \caption{Comparison of performance differences with previous studies}\label{tab:differences-with-previous-studies}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}ccc|cc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{Authors}              & \multirow{2}{*}{Model}                          & \multirow{2}{*}{Data (train, test)}         & \multicolumn{2}{c}{Performance}                           \\
            \cmidrule{4-5}
            ~                                     & ~                                               & ~                                           & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            Na, 2015~\cite{NaSH2015}              & CRF++, Lattice-based HMM                        & Sejong 200k, 50k sentences                  & 95.22                      & 97.21                        \\
            Lee et al., 2016~\cite{LeeCH2016}     & Structural SVM                                  & Sejong 666k, 74k eojeols                    & 96.41                      & -                            \\
            Li et al., 2017~\cite{Li2017}         & Seq2seq (GRU-based)                             & Sejong 90k, 10k sentences                   & 95.33                      & 97.15                        \\
            Na and Kim, 2018~\cite{NaSH2018}      & Lattice + HMM                                   & Sejong 200k, 50k sentences                  & 96.35                      & 97.74                        \\
            Min et al., 2019~\cite{MinJW2019}     & Seq2seq (Transition-based)                      & Sejong 200k, 50k sentences                  & 96.34                      & 97.68                        \\
            Song and Park, 2019~\cite{SongHJ2019} & Seq2seq (BiLSTM-based)                          & Sejong 200k, 50k sentences                  & 95.68                      & 97.43                        \\
            Youn et al, 2021~\cite{YounJY2021}    & Seq2seq (BERT-based)                            & Sejong 675k, 75k sentences                  & 95.99                      & 97.94                        \\
            Shin et al, 2023~\cite{ShinHJ2023}    & Transformer(Encoder) + BiLSTM                   & Sejong 769k, 87k sentences                  & 96.12                      & 97.74                        \\
            \midrule
            Proposed (without rerank)             & \multirow{3}{*}{Lattice + Transformer(Encoder)} & \multirow{3}{*}{Sejong 194k, 10k sentences} & 95.23                      & 97.08                        \\
            Proposed (1-stage rerank)             & ~                                               & ~                                           & 96.63                      & 97.84                        \\
            Proposed (2-stage rerank)             & ~                                               & ~                                           & \textbf{96.87}             & \textbf{98.01}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

    Korean morphological analyses have seen an influx of various methodologies in recent years~\cite{KwonHC1991, LeeDG2009, ShimKS2011, LeeJS2011, ShinJC2012, LeeCK2013, NaSH2014, NaSH2015, HwangHS2016, KimHM2016, ChungES2016, LeeCH2016, Li2017, NaSH2018, KimSW2018, ChoiYS2018, MinJW2018, MinJW2019, KimHM2019, SongHJ2019, MinJW2020, SongHJ2020, ChoiYS2020, HwangHS2020, KimHJ2021, YounJY2021, MinJW2022, KimJM2022, ShinHJ2023}.
    The nature of the Korean language, being agglutinative, introduces challenges that have propelled researchers to develop inventive solutions, many of which have laid the groundwork for future research.
    Table~\ref{tab:overview-of-recent-korean-morphological-analysis-methods} provides a concise comparison of the methodologies and key concepts of key studies that are directly or indirectly related to this research, giving a brief overview of the different methods of morphological analysis.

    \begin{table}[h!]
        \centering
        \footnotesize % OR use \tiny, \small, \footnotesize, \scriptsize
        \caption{Overview of Recent Korean Morphological Analysis Methods}
        \label{tab:overview-of-recent-korean-morphological-analysis-methods}
        \begin{tabular}{|p{10mm}|p{24mm}|p{40mm}|}
            \hline
            \textbf{Study}                        & \textbf{Methodology}                                   & \textbf{Key Concepts}                                                                                             \\
            \hline
            Na et al., 2014~\cite{NaSH2014}       & Lattice-based Discriminative Approach                  & Lattice creation from a lexicon, morpheme connectivity, path optimization in morpheme lattice, POS tagging.       \\
            \hline
            Na, 2015~\cite{NaSH2015}              & Two-stage Discriminative Approach using CRFs           & Statistical morphological analysis, CRF-based morpheme segmentation and POS tagging, full sentence application.   \\
            \hline
            Na and Kim, 2018~\cite{NaSH2018}      & Phrase-based Model with CRFs                           & Phrase-based processing units, CRF integration for morpheme segmentation and POS tagging, noise-channel modeling. \\
            \hline
            Shim, 2011~\cite{ShimKS2011}          & Syllable-based POS Tagging with CRFs                   & Syllable-based tagging, efficiency in label assignment, morphological analysis bypass.                            \\
            \hline
            Lee, 2013~\cite{LeeCK2013}            & Joint Model with Structural SVM                        & Word spacing and POS tagging joint modeling, error propagation reduction, structural SVM application.             \\
            \hline
            Lee et al., 2016~\cite{LeeCH2016}     & Hybrid Algorithm with Pre-analyzed Dictionary          & Syllable-based POS tagging, integration of pre-analyzed dictionary and machine learning, CRF application.         \\
            \hline
            Kim et al., 2016~\cite{KimHM2016}     & POS Tagging with Bi-LSTM-CRFs                          & Syllable pattern input, bi-directional LSTM and CRF for POS tagging, morpheme ambiguity handling.                 \\
            \hline
            Li et al., 2017~\cite{Li2017}         & Sequence-to-Sequence Model with Convolutional Features & Seq2seq model with convolutional features for morphological analysis, POS tagging.                                \\
            \hline
            Kim and Choi, 2018~\cite{KimSW2018}   & Integrated Model with Bidirectional LSTM-CRF           & Bidirectional LSTM and CRF for word spacing and POS tagging, syllable-based approach.                             \\
            \hline
            Choi and Lee, 2018~\cite{ChoiYS2018}  & Reranking Model with Seq2Seq Outputs                   & Seq2Seq model reranking, morpheme-unit embedding, n-gram based morpheme reordering.                               \\
            \hline
            Min et al., 2019~\cite{MinJW2019}     & Neural Transition-based Model                          & End-to-end neural transition-based learning, morpheme segmentation, sequence-to-sequence POS tagging.             \\
            \hline
            Kim et al., 2019~\cite{KimHM2019}     & Syllable Distribution Patterns with Bi-LSTM-CRF        & Utilization of syllable distribution, Bi-LSTM-CRF for morphological analysis and POS tagging.                     \\
            \hline
            Song and Park, 2019~\cite{SongHJ2019} & Tied Sequence-to-Sequence Multi-task Model             & Multi-task learning for morpheme processing and POS tagging, pointer-generator and CRF network integration.       \\
            \hline
            Song and Park, 2020~\cite{SongHJ2020} & Two-step Korean POS Tagger with Encoder-Decoder        & Encoder-decoder architecture for morpheme generation, sequence labeling for POS tagging.                          \\
            \hline
            Youn and Lee, 2021~\cite{YounJY2021}  & Two-step Deep Learning-based Pipeline Model            & Deep learning sequence-to-sequence models, BERT for morpheme restoration and POS tagging.                         \\
            \hline
            Shin and Lee, 2023~\cite{ShinHJ2023}  & Syllable-Based Multi-POSMORPH Annotation               & Syllable distribution patterns, Multi-POSMORPH tagging, Transformer encoder, BiLSTM usage.                        \\
            \hline
        \end{tabular}
    \end{table}

@@

    \subsection{Syllable-unit Morphological Analysis}\label{subsec:syllable-unit-morphological-analysis}
    To overcome the limitations of dictionary dependence, syllable-by-syllable morphological analysis has emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, Li2017, KimSW2018, ChoiYS2018, MinJW2019, KimHM2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    This method either tags each syllable and then applies a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}.
    One notable drawback is the challenge of accurately pinpointing morpheme boundaries.
    Furthermore, as the sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Recent Deep Learning Approaches}\label{subsec:recent-deep-learning-approaches}
    The incorporation of deep learning into Korean morphological analysis has brought significant advancements to the field.
    Existing deep learning methods typically employ architectures like Bidirectional Long Short-Term Memory (Bi-LSTM) networks, Convolutional Neural Networks (CNNs), and Transformer-based models.
    These approaches focus on understanding language context and sequence, utilizing the ability of these models to capture long-range dependencies and intricate patterns in text data.
    For example, Bi-LSTM-CRF models, extensively used for sequence labeling in morphological analysis, leverage LSTM's capacity to remember long-term dependencies and CRF's proficiency in sequence prediction.

    In contrast, our method innovatively integrates the re-ranking concept with BERT-based models for Korean morphological analysis.
    Unlike traditional deep learning methods that primarily use sequence-to-sequence or sequence labeling approaches, our method generates suboptimal paths using dictionary-based techniques, which are then re-ranked by BERT models.
    This dual approach, leveraging BERT's contextual understanding, allows for a more detailed and accurate morphological analysis.
    The distinction of our approach lies in its ability to address the complexities of the Korean language.
    By generating and re-ranking suboptimal paths, our method can identify and rectify anomalies that standard deep learning models may miss.
    This innovative strategy combines the precision of dictionary-based methods with the contextual comprehension of BERT models, marking a significant advancement in the field, especially for languages with intricate morphological structures like Korean.

    \subsection{Integrating Dictionary-based and Deep Learning Approaches}\label{subsec:integrating-dictionary-based-and-deep-learning-approaches}
    Tokenisation, a fundamental process in NLP deep learning models, involves breaking down text into smaller units and converting these tokens into vectors for computational processing.
    For Korean, with its complex morphological characteristics, tokenisation that respects morpheme boundaries is crucial.
    This is because it can not only accurately capture the linguistic nuances of Korean, but also improve the overall performance of deep learning models.
    This is especially important given the agglutinative nature of Korean, where words are formed by combining morphemes with different semantic and syntactic information.

    In this respect, the combination of dictionary-based morphological analysis methods and deep learning approaches used by MeCab~\cite{MeCab}, a fast and lightweight morphological analyser used for tokenisation of Korean and Japanese, is quite useful.
    The dictionary-based morphological analysis used here uses a model trained with CRFs to form a lattice structure as in ~\cite{Kudo2004, NaSH2014, NaSH2018} to identify the optimal path for morphological analysis.
    While this method provides a degree of accuracy and speed, it does not achieve the high accuracy achieved by modern deep learning.

    Our research sought to bridge this gap by effectively combining these dictionary-based morphological analysis methods with the contextual understanding capabilities of deep learning.
    Future research should continue to refine these hybrid methods to explore the possibility of end-to-end models that seamlessly combine the strengths of traditional dictionary-based analysis with the adaptive capabilities of deep learning.
    This direction promises significant advances in morphological analysis and will further push the boundaries of Korean language processing.
